On the Functionality of the Multilayer Perceptron

### 2020-12-16 Thesis/Drafts/Buell-PHYS799-ThesisDraft3.tex
Major Updates to Thesis & Figures

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being, but far more difficult for a computer to do. To execute this task, we have designed and constructed a multimodal neural network, that utilizes two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate the of the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of it's constituent unimodal architectures. Training data for this experiment is from studio recordings of the Philharmonia Symphony Orchestra and the University of Iowa's Electronic Music Studios.

@@

@@

\paragraph*{}The ability to classify time-series data into unique categories has become topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Virtanen}. In each context, the general idea is to place data or a subset of data into distinct categories called \textit{classes} based on it's properties. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to classify a purchase as fraudulent based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do unaided by some form of automation.

\paragraph*{}It has long been a goal of computer scientists and engineers alike to develop a program, algorithm, or machine that could perform a certain task to the level of human-being \cite{Bishop,Geron,Loy}. While humans are intricate organisms capable of recognition or classification in many contexts, we are not notorious for the ability to process millions of raw numerical values in the course of a few seconds. Similarly, modern computers are also complex systems that can filter through exabytes of information on a daily basis, but are not generally known for being able to distinguish cats from dogs \cite{Geron}. The compromise between humans and machines an algorithm called a \textit{neural network} that combines the computation speed of a computer with the decision-making power modeled from human brain \cite{Goodfellow,Levine,Loy}.

\paragraph*{}In this work, we have developed an automated neural network program that can classify digital sound waves as one of $37$ musical instruments that it most closely resembles. The general use of machine learning algorithms for audio information classification or identification is well explored, and a topic of much modern research \cite{Khan,Liu,Mierswa,Serizel,Zhang}. From previous work, it has been found that various types of audio classification are possible, and are still being expanded upon. Many papers details predictors for classifying speech against music or features for organizing audio-visual media based on content. In each case, features are often tailored to meet the specific task at hand \cite{Virtan,Liu}.

\paragraph*{}It has been experimentally determined that the choice of features is critically important to classification success \cite{Khan,Serizel}. In our case, the nature of these features is motivated by the physics and mechanical properties of each musical instrument class. By exploring the physics of the vibrating parts of each instrument according to the Hornbostel-Sachs classification system, we gain a more through understanding of each instrument group. This allows us to develop properties within each class that are consistent with a low variability, and what properties differ between classes and show a higher variability. These differences are key to providing a set of features which allow a neural network to differentiate between classes of musical instruments \cite{Liu,Zhang}.

\paragraph*{}As the name implies, a neural network is a computation a model or graph that is loosely structured off of the biological brain \cite{Geron,Loy}. Where a biological brain uses physical connects through axons to exchanges chemicals to communicate information, a neural network uses a chain of mathematical operations to communicate information \cite{Goodfellow}. The earliest work on attempting to model cognitive processes goes back to 1943, where Warren McCulloch and Walter Pitts began to develop a model of the human brain using principles of mathematics \cite{McCulloch,Loy}. Nearly ten years later, a simple neural networks called the \textit{perceptron} was developed by Frank Rosenblatt, which modeled connections of neurons through principles of linear algebra \cite{Geron}. Another decade after that, inspiration from the visual cortex gave rise to the convolution operation to be applied to digital image processing.

\paragraph*{}The design of our neural network function combines two architecture types by using two distinct entry arrays which are processed, and a single output is formed. This type of neural network is referred to as a \textit{multimodal} neural network \cite{Li,Ngiam}. Our model contains a convolutional neural network (CNN) that processes an image-like representation of an audio waveform, and a multilayer perceptron (MLP) that processes a list-like representation of an audio wave form. Each arm of the network is it's own computational graph that is designed to process it's respective inputs types. The activations at the end of each layer are concatenated into a new layer, which is further processed and then used to generate a single prediction. This prediction aggregated from two features on average shows superior classification when compared to the performance either of either of it's constituent unimodal architectures alone.

@@

\paragraph*{}This project includes a number of steps to ensure it's completion. Here we show a high-level outline of the pieces involved.

@@

\paragraph*{}Consider the biological process of hearing a sound wave and matching it to a source. We can model this behavior by some unknown function $F$ and approximate it with the function $F^*$. This can map the contents of a sound file to a potential source as to mimic the behavior of $F$. For a set of inputs (called features or predictors) $\vec{x} = \big\{ x_0 , x_1 , x_2 , ... , x_{p-1} \big\}$ and a set of classes $0$ through $k-1$, we denote this function as:
\begin{equation}
\label{eqn-MappingFunction}
F^*: \vec{x} \rightarrow \big\{ 0 , 1 , 2 , ... , k-1 \big\}
\end{equation}

\paragraph*{}In secs. (\ref{subsec-NerualNetworkIntro}) and (\ref{subsec-NetworkStructure}), we introduce the nature and structure of a neural network and how it can use used in part to solve this problem. We discuss it's inspiration from a biological brain,and show how different layer types are influenced from the brain's functions as well as how they are used in this problem.

@@

\paragraph*{}In order to train the neural network to be able to identify musical instrument sources, we need a suitable data set to present to the model as training data. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files (See acknowledgments). These files contain short segments of musical instruments performing a single note, and is labeled according to the instrument that produced the sound. The files and labels make up the core of the data set that we use to train the model.

\paragraph*{}To ensure that these data sets are formatted consistently, we read each sample from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in section (\ref{subsubsec-Preprocessing}).

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of predictors or features which represent the raw input given to the model. These inputs are designed to represent properties of wave forms that can be expressed by a single or multiple numerical values. We detail these features in section (\ref{sec-Features}). These are used in place of a full waveform to represent a file's contents in a low-dimensional array. We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of output categories determined, we can organize the structure of the neural network function. The shape of the network defines the flow of information, and the hypothesis space. The nature of all possible solutions that the trained model could arrive at is subject to the architecture \cite{Goodfellow,Loy}. We construct a multimodal neural network, explored in section (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to evaluate each input independently, and concatenate the results to produce a single outputted prediction.

@@

\paragraph*{}We execute a K-Folds cross validation algorithm to evaluate the performance of the neural network across overlapping subsets of data. This ensures that the model can perform consistently and reliably given small changes in initial conditions. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore the explore this is section (\ref{sec-PerfEval}).

\subsubsection{Running Predictions of Chaotic Synthesizer Files}

\paragraph*{}Once we have established a suitable performance of the model, we allow the model to run predictions on the un-labeled chaotic synthesizer wave forms. We output the prediction results to a file, and compare the neural network predictions against human predictions. If further corrections are needed, we revert and re-deign the features, architecture, or hyper-parameters as needed. We explore thisas part of our conclusions in section (\ref{sec-Conclusion}).

@@

\paragraph*{}The task of automatically matching sound files to musical instrument classes is exceedingly challenging through conventional computer programming techniques. Because of the complexity of sound recognition, the challenge arises as to how to build some sort of program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for a classification program, we seek an solution that allows for a computer to \textit{learn} from labeled training examples in order \textit{teach} itself \cite{Bishop,Mitchell}. We choose to employ a neural network, which a computer algorithm that is designed to update itself to perform a task with increasing proficiency as it is presented with new information \cite{Geron2,Goodfellow,Levine,Mitchell}.

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro, and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result is a computer program that is reminiscent of the a brain in that it structured much like the brain, and can\textit{learn} from new data. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Levine,Bishop}. Where a brain uses physical connection of neurons through axons, the neural network uses mathematical connections of neurons through mathematical operations \cite{Geron,Levine}.

@@

\paragraph*{}A Neural Network is a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes an input, typically an array of real double-precision floating-point numbers, and returns an output which is a modified array of new double-precision floating-point numbers. The exact nature of this operation can be very different depending on the layer type or where it sits within the network. Typically, a neural network is characterized by a chain-like structure where the output of one layer is fed directly into the input of the next layer \cite{Geron,Goodfellow}. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. The output of the final layer encodes the "decision" of the model given a particular input.

@@

\paragraph*{}Classification tasks require a set of inputs called \textit{predictors} or \textit{features} \cite{James,Loy,Serizel}. A feature is a quantitative, low-dimensional representation of a sample that conveys its important characteristics. For example, in classifying types of animals, a feature could be the mass, volume or number of teeth on the animal. Typically, we produce a set of features $p$ for each sample and assemble them into a $p$-dimensional vector called a \textit{feature vector} which acts as a list of descriptive qualities of the sample \cite{Geron,James}.
Audio machine learning researchers M. Kashif Saeed Khan and Wasfi G. Al-Khatib describe the vitality of feature selection \cite{Khan}:
\begin{quote}
The data reduction stage which is also called feature extraction, consists of discovering a few important facts about each class. The choice of features is critical as it greatly affects the accuracy of audio classification. 
\end{quote}
An adequate selection of features is vital to the process of training and evaluation of the classifier model \cite{Mierswa,Serizel,Liu}. Consider the task of identifying fish from sharks using only the presence of gills. Since both sharks and fish have gills, the features are correct representations, but not useful for discerning them. A more appropriate choice would be compare the number of gills that fish or sharks have, which would provide a correct representation and allows for differentiation.

@@

\paragraph*{}Window functions are use to account for discontinuities that may arise in the waveform at the edges of each analysis frame. The Fourier Transform assumes that an integer number of signal periods fit within each analysis frame, however this is rarely the case. Transforming a signal with to \textit{spectral leakage} where energy appears to "leak" into adjacent frequency bins. This result is a frequency domain that is not an accurate representation of the equivalent time domain. The Hanning window uses the raised cosine function to taper the signal off to zero at the end-points, thus eliminate discontinuities at the edges while weighting the samples at the center much more heavily. This allows for a much cleaner transform array.

@@

\paragraph*{}Despite the higher average scores in the hybrid network, see Fig. (\ref{fig-MultimodalXval}) compared to each unimodal network, Fig. (\ref{fig-UnimodalXVal}), some exceptions do arise at a micro-level (within each class). Consider the Hi-hat instrument in Fig. (\ref{fig-PercussionScores}c). over the cross validations, it was found that the multilayer perceptron architecture produced a better precision, recall, and F1 score than either the convolution or the hybrid network architecture. This can be contrasted with most other classes showing hat the hyrbidg model performed better than either individual mode, with the convolution architecture as a close second.

\paragraph*{}We can also observe a loose correlation to the number of samples in each class compared to the classification metric scores. When a model is trained on one class more than another we are more likely to generate predictions in that class too. This can be seen in the scores and occurrences of alto flutes (\ref{fig-HighWindsScores}a) with $72$ samples compared to that of English horns (\ref{fig-HighWindsScores}e) with $1382$ samples. In both confusion matrices of Fig. (\ref{fig-MultimodalConfs}), we see a realty small entry in $C_{0,0}$ (alto flutes) compared to $C_{14,14}$ ( English horn). This indicates that fewer correction predictions were made, and the normalized matrix shows that fewer predictions were made at all. Contrast this with the english horn having a dark red square in both matrices shows many predictions, and many correction predictions.

\paragraph*{}
\textcolor{red}{Also not done here}

@@

\paragraph*{}Examining the two diagonally dominant confusion matrices in Fig. (\ref{fig-MultimodalConfs}) indicates that the neural network is consistently making correct categorical predictions. The presence of off-diagonal entries in the confusion matrices indicates that misclassifications are a possibility due to an imperfect set features. Averaged across $37$ classes, we see in Fig. (\ref{fig-MultimodalXval}) that the classification accuracy score averages around $0.98$ and precision and recall scores around $0.83$ and $0.82$ respectively. On average, examining each class in an \textit{one-versus-all} context, the model selects $83\%$ of relevant samples, and $82\%$ of relevant samples are selected. This indicates a neural network that can reasonably differentiate between $37$ classes of musical instruments.

\paragraph*{}Additionally, we conclude that on average, the hybrid neural network architecture in Fig. (\ref{fig-NetworkArchitecture}) exhibits superior classification performance over either of it's unimodal counterpart architectures. In examining confusion matrices from both unimodal models in Fig. (\ref{fig-UnimodalConfs}), we see the presence of diagonal dominance. However, in comparing this to the multimodal network confusion matrix from Fig. (\ref{fig-MultimodalConfs}), we see that the multimodal network has a much stronger main diagonal entry. This indicates that my aggregating to the output of each modal branch with a concatenation layer, we can benefit from the predictive power of both branches together. This indicates that both a multilayer perceptron and convolutional network are producing generally appropriate classification labels, but an aggregation of two architectures yields and even higher classification performance.

\text{red}{Also plz help here}

@@

\item[•]Dakota Buell, Tan Dao, John Parker, Nathan Richard, and Morgan Saidel for consultation on physics, mathematics, and programming topics.

\item[•]Madeline Edwards for differing me to this project, as well as constant support and consultation on mathematical topics.

@@

\bibitme{Matplotlib}
John D. Hunter. Matplotlib: A 2D Graphics Environment, Computing in Science & Engineering, 9, 90-95 (2007), DOI:10.1109/MCSE.2007.55

@@

@@

>>>>

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being in many cases, but is far more difficult for a computer to do. In order to automate source source identification, we have designed and constructed a multimodal neural network that can classify audio files by what musical instrument the sound most resembles. The network uses two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate and transform the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of it's constituent unimodal architectures. Training data for this experiment is from studio recordings of the Philharmonia Symphony Orchestra and the University of Iowa's Electronic Music Studios.

@@

\vspace{8cm}

\begin{center}
All source code and documentation for this project can be found at  \\
\textit{https://github.com/landonbuell/Buell-Senior-Thesis}
\end{center}

@@

\paragraph*{}The ability to classify time-series data into unique categories has become topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Virtanen}. In each context, the general idea is to place information into distinct categories called \textit{classes} based on inherent properties of the data. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to classify a purchase as fraudulent based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do without the aide of some form of automation.

\paragraph*{}It has long been a goal of computer scientists and engineers alike to develop a program, algorithm, or machine that could perform a certain task to the level of human-being \cite{Bishop,Geron,Loy}. While humans are intricate organisms capable of recognition or classification in many contexts, we do not have the ability to process millions of raw numerical values in the course of a few seconds. Similarly, modern computers are also complex systems that can filter through exabytes of information on a daily basis, but are not generally known for being able to distinguish cats from dogs \cite{Geron}. The compromise between humans and machines an algorithm called a \textit{neural network} that combines the computation speed of a computer with the decision-making power modeled from human brain \cite{Goodfellow,Levine,Loy}.

\paragraph*{}In this work, we have developed an automated neural network program that can classify digital sound waves as one of $37$ musical instruments that it most closely resembles. The general use of machine learning algorithms for audio information classification or identification is well explored, and a topic of much modern research \cite{Khan,Liu,Mierswa,Serizel,Zhang}. From previous work, it has been found that various types of audio classification are possible, and are still being expanded upon. Many papers detail possible sets of inputs for classifying speech against music or inputs for organizing audio-visual media based on content. In each case these inputs, called features, must often be tailored to meet the specific task at hand \cite{Virtanen,Liu}.

\paragraph*{}It has been experimentally determined that the choice of features is critically important to classification success \cite{Khan,Mierswa,Serizel}. In our case, the nature of these features is motivated by the physics and mechanical properties of each musical instrument class. By exploring the physics of the vibrating parts of each instrument according to the Hornbostel-Sachs classification system, we gain a more thorough understanding of each instrument group. This allows us to develop properties within each class that are consistent and have low variability, and what properties differ between classes and show a higher variability. These differences are key to providing a set of features which allow a neural network to differentiate between classes of musical instruments \cite{Liu,Zhang}. For our classifier model, we have developed $24$ features from the time-series and frequency-series representation of the audio file data. These quantities are assembled into a single axis array for processing. Additionally, we construct an image-like representation of the waveform and the 2D array is also used for proceessing. Both of these input modes represent a set of features that convey the nature of the waveform in two different modalities.

\paragraph*{}As the name implies, a neural network is a computation a model or graph that is loosely structured off of the biological brain \cite{Geron,Loy}. Where a biological brain uses physical connects through axons to exchanges chemicals to communicate information, a neural network uses a chain of mathematical operations to communicate information \cite{Goodfellow}. The earliest work on attempting to model cognitive processes goes back to 1943, where Warren McCulloch and Walter Pitts began to develop a model of the human brain using principles of mathematics \cite{McCulloch,Loy}. Nearly ten years later, a simple neural networks called the \textit{perceptron} was developed by Frank Rosenblatt, which modeled connections of neurons through principles of linear algebra \cite{Geron,Petrik}. Another decade after that, inspiration from the visual cortex gave rise to the convolution operation to be applied to digital image processing.

\paragraph*{}The design of our neural network function combines two distinct architecture types by using two distinct entry arrays which are processed, and a single output is formed. This type of neural network is referred to as a \textit{multimodal} neural network \cite{Li,Ngiam}. Our model contains a convolutional neural network (CNN) that processes an image-like representation of an audio waveform, and a multilayer perceptron (MLP) that processes a list-like representation of an audio wave form. Each arm of the network is it's own computational graph that is designed to process it's respective inputs types. The activations at the end of each layer are concatenated into a new layer, which is further processed and then used to generate a single prediction. This output that is aggregated from two inputs shows on average, superior classification performance when compared to either of it's constituent unimodal architectures alone.

@@

\paragraph*{}This project includes a number of steps from taking raw input data to the final classification task. In this section, we provide a very high level outline of the parts of this projects and show how each one influence the final product. Additionally, we indicate the sections in this thesis where the particular material is introduced and developed much more thoroughly.

@@

\paragraph*{}To automate the process of classifying audio files, we require a program or algorithm that can make practical decisions based on a prescribed set of information. This is to say that we need a \textit{function} to process an input or set of inputs, and return a predicted label that corresponds to a musical instrument. Consider the biological process of hearing a sound wave and matching it to a source. We can model this behavior by some unknown function $F$ and approximate it with the function $F^*$. This approximate function is composed of smaller functions that collectively work together to direct the flow of information and allow for the mapping of the contents of a sound file to a potential source as to mimic the behavior of $F$. For a set of inputs $\vec{x} = \big\{ x_0 , x_1 , x_2 , ... , x_{p-1} \big\}$ and a set of classes $0$ through $k-1$, we denote this function as:
\begin{equation}
\label{eqn-MappingFunction}
F^*: \vec{x} \rightarrow \big\{ 0, 1, 2, ... ,k-2, k-1 \big\}
\end{equation}

\paragraph*{}The smaller functions that comprise $F^*$ are called \textit{layers} \cite{Loy,Goodfellow}. Each type of layer performs a different operation and processes different types of input. In secs. (\ref{subsec-NerualNetworkIntro}) and (\ref{subsec-NetworkStructure}), we introduce the nature and structure of a neural network and how it can use used in part to solve this problem. We discuss it is inspired from a biological brain and show how different layer types are influenced from the brain's functions as well as how they are used in this problem. Using information about the problem, we carefully design the inputs, outputs and structure of the neural network function to be able to complete this task in a practical manner.

@@

\paragraph*{}In order to train the neural network to be able to identify musical instrument sources, we need a suitable data set to present to the model as training and validation data. The models uses this data and corresponding labels in the training process to modify it's function to allow for it to complete the task. University of Iowa Electronic Music Studio, and the London-Based Philharmonia Orchestra each have a large collection of publicly available audio files \cite{Philharmonia,UnivIowa}. These files contain short segments of musical instruments performing a single note, and are labeled according to the instrument that produced the sound. The files and labels make up the core of the data set that we use to train the model.

\paragraph*{}To ensure that these data sets are formatted consistently, we have designed a MATLAB program to read each audio file from it's original format, \textit{.AIF}, \textit{.MP3}, or similar, and rewrite the data as a new \textit{.WAV} files, sampled at 44.1 kHz with a 24-bit depth. This ensures that all data will have a consistent format when features are extracted. This stage is detailed in section (\ref{subsubsec-Preprocessing}). After the samples are formatted, we have produced a Python script with iterates through each new file and assigns a class label based on the title of the instrument. This has been design to account for the specific naming cases that arise. We assemble a dictionary-like structure of all training file with organizes the local file location of each sample and provides an integer and string label for all samples.

@@

\paragraph*{}The performance of a neural network is largely dependent of designing an appropriate set of predictors or features which represent the raw input given to the model. These inputs are designed to represent prolific properties of wave forms that can be expressed by a single or multiple numerical values. We detail these features in section (\ref{sec-Features}). We use tools from physics, mathematics, and signal processing to define and explore a comprehensive set of features that enables a high performance of the classifier.

@@

\paragraph*{}With the appropriate set of features designed, and the number of output categories determined, we can organize the structure of the neural network function. The shape of the network defines the flow of information, and the hypothesis space. The nature of all possible solutions that the trained model could arrive at is subject to the architecture \cite{Goodfellow,Loy}. We construct a multimodal neural network, explored in section (\ref{subsec-Architecture}), that processes two distinct entry points derived from the same audio sample. This network is designed to evaluate each input independently, and concatenate the results to produce a single outputted prediction. Out nueral network was implemented using the tools available with the \textit{Tensorflow} module \cite{Tensorflow}.

@@

\paragraph*{}We execute a K-Folds cross validation algorithm to evaluate the performance of the neural network across overlapping subsets of data. This ensures that the model can perform consistently and reliably given small changes in initial conditions. In this stage, we also choose the value of hyper-parameters, activation functions and layer widths to best compliment the chosen features. This process is repeated and expanded upon until we have produced a model with a sufficient performance. We explore the explore this is section (\ref{sec-PerfEval}). Results of all metrics are computed with the \textit{numpy},\textit{scipy}, and \textit{scikit learn} python modules \cite{numpy,scipy,sklearn}. All performance evaluate visualiztions are produced with the matplotlib python package \cite{matplotlib}.

@@

\paragraph*{} Because of the complexity of sound recognition, the challenge arises as to how to build some sort of program that could function at a level above general procedural or explicit instructional rules. Rather than \textit{hard-coding} a set of conditions or parameters for a classification program, we seek an solution that allows for a computer to \textit{learn} from labeled training examples in order \textit{teach} itself \cite{Bishop,Mitchell}. We choose to employ a neural network, which a computer algorithm that is designed to update itself to perform a task with increasing proficiency as it is presented with new information \cite{Geron2,Goodfellow,Levine,Mitchell}.

\paragraph*{}As the name implies, early neural networks were inspired from from the human brain. Former YouTube video classification team lead, and current machine learning consultant, Aurelien Geron writes about the relationship between biological brains and mathematical neural networks \cite{Geron}: 
\begin{quote}
Birds inspired us to fly, burdock plants inspired velcro, and nature has inspired many other inventions. It seems only logical, then, to look to the brain's architecture for inspiration on how to build an intelligent machine.
\end{quote}
The result is a computer program that is reminiscent of the a brain in that it structured much like the brain, and can\textit{learn} from new data. We model operations with collections are artificial neurons that are capable of connect and interacting with other neurons \cite{Levine}. Where a biological brain uses connections of chemical and electrical exchanges to process input and make decisions, the computational counterpart uses arithmetic and numerical exchanges to similarly process input and make decisions \cite{Levine,Bishop}. Where a brain uses physical connection of neurons through axons, the neural network uses mathematical connections of neurons through mathematical operations \cite{Geron,Levine}. We call the numerical value contained within an artificial neuron the \textit{activation} of that neuron \cite{Goodfellow,Geron}.

@@

\paragraph*{}A Neural Network is a model of a mathematical function, composed of several smaller mathematical functions called \textit{layers} \cite{Goodfellow,Loy}. Each layer represents an operation that takes an input, and returns an output . The exact nature of this operation can be very different depending on the layer type or where it sits within the network. Typically, a neural network is characterized by a chain-like structure where the output of one layer is fed directly into the input of the next layer \cite{Geron,Goodfellow}. It is this process of transforming inputs successively in a particular order until an output is attained that makes up the core functionality of a neural network \cite{Geron,Loy}. The output of the final layer encodes the "decision" of the model given a particular input.

@@

\paragraph*{}Classification tasks require a set of inputs called \textit{predictors} or \textit{features} \cite{James,Loy,Serizel}. A feature is a quantitative, low-dimensional representation of a sample that conveys its important characteristics. For example, in classifying types of animals, a feature could be the mass, volume or number of teeth on the animal. Typically, we produce a set of $p$ features for each sample and assemble them into a $p$-dimensional vector called a \textit{feature vector} which acts as a list of descriptive qualities of the sample \cite{Geron,James}. This list is presented to the neural network and used to 
Audio machine learning researchers M. Kashif Saeed Khan and Wasfi G. Al-Khatib describe the vitality of feature selection \cite{Khan}:
\begin{quote}
The data reduction stage which is also called feature extraction, consists of discovering a few important facts about each class. The choice of features is critical as it greatly affects the accuracy of audio classification. 
\end{quote}
An adequate selection of features is vital to the process of training and evaluation of the classifier model \cite{Mierswa,Serizel,Liu}. Consider the task of identifying fish from sharks using only the presence of gills. Since both sharks and fish have gills, the features are correct representations, but not useful for discerning them. A more appropriate choice would be compare the number of gills that fish or sharks have, which would provide a correct representation and allows for differentiation.

@@

\paragraph*{}Window functions are use to account for discontinuities that may arise in the waveform at the edges of each analysis frame. The Fourier Transform assumes that an integer number of signal periods fit within each analysis frame, however this is rarely the case. Transforming a signal with to \textit{spectral leakage} where energy appears to "leak" into adjacent frequency bins. This result is a frequency domain that is not an accurate representation of the equivalent time domain. The Hanning window uses the raised cosine function to taper the signal off to zero at the end-points, thus eliminate discontinuities at the edges while weighting the samples at the center much more heavily. This allows for a much cleaner transform into frequency space.

@@

\paragraph*{}Despite the higher average scores in the hybrid network, see Fig. (\ref{fig-MultimodalXval}) compared to each unimodal network, Fig. (\ref{fig-UnimodalXVal}), some exceptions do arise at a micro-level (within each class). Consider the Hi-hat instrument in Fig. (\ref{fig-PercussionScores}c). Over the cross validations, it was found that the multilayer perceptron architecture produced a better precision, recall, and F1 score than either the convolution or the hybrid network architecture.

 This can be contrasted with most other classes showing hat the hybrid model performed better than either individual mode, with the convolution architecture as a close second.

\paragraph*{}While the hybrid model can differentiate between many types of instruments, each uni-modal variant shows slightly less success over the full data set. In many instances, the MLP-only architecture particularly shows unstable and inconsistent performance. For certain instruments like the $E\flat$ clarinet, tenor saxophone, tenor trombone, and hi-hat cymbal, the feature vector does a very poor job of isolating and identifying those classes.  However, for instruments such as the soprano saxophone, tuba, and square wave the feature-vector allows the MLP to perform very reasonably. We see a similar imbalance with the CNN-only architecture where classes such as alto flute, bass flute, bass trombone, and mandolin show poor classification rates. This can be contrasted with English horn, tuba, bells, and double bass, where a much higher classification performance is found. Further exploration is required to provide a full explanation and justification of why this wide range of performance is observed.

\paragraph*{}We can also observe a loose correlation to the number of samples in each class compared to the classification metric scores. When a model is trained on one class more than another we are more likely to generate predictions in that class too. This can be seen in the scores and occurrences of alto flutes (\ref{fig-HighWindsScores}a) with $72$ samples compared to that of English horns (\ref{fig-HighWindsScores}e) with $1382$ samples. In both confusion matrices of Fig. (\ref{fig-MultimodalConfs}), we see a relatively small entry in $C_{0,0}$ (alto flutes) compared to $C_{14,14}$ ( English horn). This indicates that fewer correction predictions were made, and the normalized matrix shows that fewer predictions were made at all. Contrast this with the English horn having a dark red square in both matrices shows many predictions, and many correction predictions.

\paragraph*{}Future work on this project would benefit from a further exploration and development of input features. While the ability to discern classes is present, classification scores are still unstable and we observe a wide range of performances. To combat this, we would produce additional feature that convey characteristic properties of the digital audio files that allow for the generation of more distinctive decision boundaries. In addition to new features, it would be beneficial to further study the distribution of samples in feature space and execute some form of $K$-best features algorithm. This would indicate which are the chosen predictors show properties that are inconsistency with the ideas of intra-class and inter-class variability.

\paragraph*{}In addition to input features, it may be beneficial to group for the the related classes together. For example, violins and violas have very similar physic construction, have very similar formant structures, and thus have very similar behaviors in our feature-space. As a result, some humans may also have difficulty differentiating the two, so it may be forgiven that a classification neural network would do the same. We can also consider grouping range variants of parent instruments together such as flutes, alto flutes, bass flutes into a single \textit{flutes} category. It is possible that the new distribution of properties in feature-space may also warrant the development of a new set of predictors again. It would also be worth exploring how classification performance would be affected when weighting each class inversely to it's number of samples in the data set. This would allow each class to be weighted uniformly in the training process and may potentially mitigate some of the variability of the performances across classes.

\paragraph*{}Finally as a practical demonstration, future work entails the deployment of this classifier to generally unlabeled audio samples, provided that they have the same format. In particular, we would like to explore how classification would behave when testing it on digitally generated chaotic synthesizer wave forms. These signals are representations of chaotic systems that resemble some of the acoustic the periodic properties of acoustic waveforms. By suing the automated classifier, we would like to test the versatility of the classifier when subject to samples that may differ drastically in same ways from the training data set. This demonstration would be a fantastic exploration into the ability for the model to generalize, and the strength of the learned function parameters.

@@

\paragraph*{}Examining the two diagonally dominant confusion matrices in Fig. (\ref{fig-MultimodalConfs}) indicates that the neural network is consistently making correct categorical predictions. The presence of off-diagonal entries in the confusion matrices indicates that misclassifications are a common possibility. Averaged across $37$ classes, we see in Fig. (\ref{fig-MultimodalXval}) that the classification accuracy score averages around $0.98$ and precision and recall scores around $0.83$ and $0.82$ respectively. This means that examining each class in an \textit{one-versus-all} context, the model selects $83\%$ of relevant samples, and $82\%$ of relevant samples are selected. This indicates that the multimodal neural network that can reasonably differentiate between $37$ classes of musical instruments.

\paragraph*{}We conclude that on average, the hybrid neural network architecture in Fig. (\ref{fig-NetworkArchitecture}) exhibits superior classification performance over either of it's unimodal counterpart architectures. In examining confusion matrices from both unimodal models in Fig. (\ref{fig-UnimodalConfs}), we see the presence of diagonal dominance. However, in comparing this to the multimodal network confusion matrix from Fig. (\ref{fig-MultimodalConfs}), we see that the multimodal network has a much stronger main diagonal entry. This indicates that my aggregating to the output of each modal branch with a concatenation layer, we can benefit from the predictive power of both branches together. This indicates that both a multilayer perceptron and convolutional network are producing generally appropriate classification labels, but an aggregation of two architectures yields and even higher classification performance.

@@

\item[•]Dakota Buell, Tan Dao, Nathan Richard, and Morgan Saidel for consultation on physics, mathematics, and programming topics.

\item[•]Madeline Edwards and John Parker for differing me to this project, as well as constant support and consultation on mathematical topics.

@@

\bibitem{Matplotlib}
John D. Hunter. Matplotlib: A 2D Graphics Environment, Computing in Science \& Engineering, 9, 90-95 (2007), DOI:10.1109/MCSE.2007.55

@@

\bibitem{Philharmonia}
Philharmonia Symphony Orchestra home page- \textit{https://philharmonia.co.uk/}

@@

\bibitem{UnivIowa}
University of Iowa Electronic Music Studios home page- \textit{http://theremin.music.uiowa.edu/}
